### 2022-09-28 chapters/optimization/python-dir/python.tex
add : conculsion

In this chapter, we aim to reduce the energy consumption of Python code.
We first start by presenting the motivation behind our work in Section~\ref{python:sec_motivation}.
In the second section, we share some insights on the impact of programmer choices. To do this, we will run a series of experiments to examine the energy behavior of Python code in the most of its use cases.Section\ref{python:sec_insights} will then end by investigating several features of Python structure and their impact on energy consumption.
After that, in Section~\ref{python:sec_interpreters},we investigate other non-intrusive approaches to optimizing the energy consumption of applications by comparing multiple Python implementations, which include alternative interpreters and libraries that are dedicated to optimizing the code without changing its structures, such as ahead-of-time (AOT) compilation and just-in-time (JIT) libraries that are maintained by the community.

@@

\subsection{Python Gluttony}
% In the other hand,because of the interpreted nature of Python code, this programming language has paid a high price in terms of performance~\ref{fig:clbg}, memory and energy consumption.
According to~\cite{pinto_energy_2017} and~\cite{noureddine_preliminary_2012}, Python tends to be one more energy hungry programming language.
As one can notice in \Cref{fig:hannoi}, Python consumes $30$ times more than C or C++.
The benchmark was done with an implementation of the \fnurl{Tower of Hanoi}{https://en.wikipedia.org/wiki/Tower_of_Hanoi} of 30 disks.

@@

Python consumes energy, mainly because it is slow in execution.
Its flexibility and simplicity caused it to drop off in performance because Python gains its flexibility from being a dynamic language.
Therefore, it requires a faster interpreter to execute its programs to compete against alternatives written in compiled programming languages, such as C and C++, or semi-compiled languages like Java.

@@

cref{fig:eopochvsaccuracy}shows the accuracy of the model based on the number of epochs. This figure shows that there is a non-significant impact on the choice of the strategy on the accuracy; it is only a matter of the number of epochs. Hover, the increase in accuracy is not linear, as can be seen. It took only 10 epochs to reach an accuracy of 84\%, but it required more than twice the number of epochs to add axtra 6\% accuracy. This highlights the price that one should pay to increase the accuracy of the model. This has been confirmed by the energy consumption, as shown in \cref{fig:cum_energy_vs_accuracy}.
As one can see the training the model up to 90\% accuracy requires three times the energy required for training to 80\% accuracy. Moreover, one can notice that this price is mostly payed by the calcul units such as CPU and GPU, while the memory is not that impacted.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/epoch_duration.pdf}
    \caption{Distribtion of the duration of each epoch.}
    \label{fig:epoch_duration}
\end{figure}
\subsubsection{conculsion}
We discovered that when the model's accuracy improves, so does the energy required for a subsequent accuracy gain.
This begs the question of when we should discontinue training.
Is a 10\% increase in accuracy worth it if we have to spend three times the energy?
While using the GPU, the CPU has no significant impact on the energy consumption of the training. Therefore, one can limit the number of CPU cores used during the execution without sacrificing either the execution time nor the accuracy.

@@

Finally, we run the same experiment with the \textsf{Optimized} strategy using JouleHunter.
Figure~\ref{fig:django_profiled_optimized} depicts the resulting energy consumption.
As one can notice, while the rendering method consumed the same amount of energy as in the \textsf{Vanilla} strategy (around 1.3\,kJ), the resolve part dropped by $20\times$.

@@

\paragraph*{Synthesis}
This study demonstrates that the optimal way to reduce the energy consumption of Python application is to follow the guidelines and to privilege the built-in functions.

@@

\paragraph*{Multiprocessing}
Unlike the multi-threading strategy, the multi-processing strategy is based on forking the program into multiple processes, which are independent from each other. This allows the Python interpreter to run multiple processes at the same time, which explains the better performance of the multi-processing strategy. Howerver, this will increase the Average power consumption of the CPU, which is around $40 Watts$ compared to the sequential one, which is around $22 Watts$. nevertheless, the multi-processing strategy is still the most energy efficient one.

Despite, the fact that increasing the number of processes will reduce the execution time because we split the tasks on the number of processes, there is a point where the execution time will increase. As one can see in \Cref{fig:python_multiprocessing}, the energy consumption is correlated with the number of processes until reaching the limit of physical cores, when concurrent processes start to compete for the CPU.

@@

\paragraph*{Synthesis}
This study demonstrates that the optimal way to reduce the energy consumption of Python applications is to use the multiprocessing strategy, which is the most energy efficient one. However, this strategy should be used with caution as it can lead to increased execution times.
This study demonstrates that the optimal number of processes is equal to the number of physical cores, which is four in our case.
By using the multithreading strategy, it was also found that sometimes the performance of the application has to be given up in order to save energy.

\clearpage
\input{python_interperters.tex}
% \subsection{threads to validity \note{missing}}
\clearpage
\section{Summary}
Because Python is widely used by software services deployed in public and private cloud infrastructures, increasing the energy efficiency of Python-based applications will reduce ICT carbon emissions.
This chapter discusses various ways for increasing the energy efficiency of Python-based applications. It first explains why such a choice was made, and then it develops a method to enhance the energy efficiency of the most commonly employed use cases, which are revealed to include web development and machine learning.
In web development, we found that the majority of the energy is wasted while processing requests, therefore adopting the proper strategy can reduce energy consumption by tenfold.
Following that, we looked at the energy usage for the machine learning usecase and discovered that it is not proportional to model accuracy.
As a result, a reduction in accuracy can result in large energy savings.
Following that, we dug deeper to investigate how data structures, parallelism, and iterative methods affect energy use.
In the second section, we presented a simple technique for optimizing the energy usage of Python-based apps. without making substantial changes to the code This is the use of many Python interpreters
Section 2 started by categorizing and filtering these implementations into three major kinds (compiler, interpreter, and extra libraries).
Then we conducted a series of experiments to compare the energy consumption of various interpreters.
The findings indicate the lack of a general solution and the importance of selecting the appropriate interpreter for the appropriate use case. In certain circumstances, some of these answers may be the best, while in others, they may be the worst.
Finally, regardless of implementation, we demonstrated that using JIT is the most efficient technique to boost the energy efficiency of Python-based programs.
We believe that our contributions will benefit a wide spectrum of legacy systems, not only reducing ICT's carbon footprint, but also lowering their cloud bills for the resources these services need.

>>>>

In this chapter, we aim to reduce the energy consumption of Python code.
We first start by presenting the motivation behind our work in Section~\ref{python:sec_motivation}.
In the second section, we share some insights on the impact of programmer choices. To do this, we will run a series of experiments to examine the energy behavior of Python code in most of its use cases.Section\ref{python:sec_insights} will then end by investigating several features of Python structure and their impact on energy consumption.
After that, in Section~\ref{python:sec_interpreters}, we investigate other non-intrusive approaches to optimizing the energy consumption of applications by comparing multiple Python implementations, which include alternative interpreters and libraries that are dedicated to optimizing the code without changing its structures, such as ahead-of-time (AOT) compilation and just-in-time (JIT) libraries that are maintained by the community.

@@

\subsection{Python Gluttony}
% On the other hand, because of the interpreted nature of Python code, this programming language has paid a high price in terms of performance~\ref{fig:clbg}, memory and energy consumption.
According to~\cite{pinto_energy_2017} and~\cite{noureddine_preliminary_2012}, Python tends to be one more energy hungry programming language.
As one can notice in \Cref{fig:hannoi}, Python consumes $30$ times more than C or C++.
The benchmark was done with an implementation of the \fnurl{Tower of Hanoi}{https://en.wikipedia.org/wiki/Tower_of_Hanoi} of 30 disks.

@@

Python consumes energy mainly because it is slow in execution.
Its flexibility and simplicity caused it to drop off in performance because Python gains its flexibility from being a dynamic language.
Therefore, it requires a faster interpreter to execute its programs to compete against alternatives written in compiled programming languages, such as C and C++, or semi-compiled languages like Java.

@@

cref{fig:eopochvsaccuracy}shows the model's accuracy based on the number of epochs. This figure shows that there is a non-significant impact on the choice of the strategy on the accuracy; it is only a matter of the number of epochs. Hover, the increase in accuracy is not linear, as can be seen. It took only ten epochs to reach an accuracy of 84\%, but it required more than twice the number of epochs to add axtra 6\% accuracy. This highlights the price that one should pay to increase the accuracy of the model. This has been confirmed by the energy consumption, as shown in \cref{fig:cum_energy_vs_accuracy}.
As one can see the training the model up to 90\% accuracy requires three times the energy required for training to 80\% accuracy. Moreover, one can notice that this price is mostly payed by the calcul units such as CPU and GPU, while the memory is not that impacted.

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{imgs/epoch_duration.pdf}
    \caption{Distribtion of the duration of each epoch.}
    \label{fig:epoch_duration}
\end{figure}
\subsubsection{Synthesis}
We discovered that when the model's accuracy improves, so does the energy required for a subsequent accuracy gain.
This begs the question of when we should discontinue training.
Is a 10\% increase in accuracy worth it if we have to spend three times the energy?
While using the GPU, the CPU has no significant impact on the energy consumption of the training. Therefore, one can limit the number of CPU cores used during the execution without sacrificing either the execution time nor the accuracy.

@@

Finally, we run the same experiment with the \textsf{Optimized} strategy using JouleHunter.
Figure~\ref{fig:django_profiled_optimized} depicts the resulting energy consumption.
As one can notice, while the rendering method consumed the same amount of energy as in the \textsf{Vanilla} strategy (around 1.3\,kJ), the resolve part dropped by $20\times$.
\paragraph{Synthesis}

@@

\paragraph*{Synthesis}
This study demonstrates that the optimal way to reduce the energy consumption of Python applications is to follow the guidelines and to privilege the built-in functions.

@@

\paragraph*{Multiprocessing}
Unlike the multi-threading strategy, the multi-processing strategy is based on forking the program into multiple processes, which are independent of each other. This allows the Python interpreter to run multiple processes simultaneously, which explains the better performance of the multi-processing strategy. However, this will increase the Average power consumption of the CPU, which is around $40 Watts$, compared to the sequential one, which is around $22 Watts$. Nevertheless, the multi-processing strategy is still the most energy-efficient one.

Although increasing the number of processes will reduce the execution time because we split the tasks on the number of processes, there is a point where the execution time will increase. As one can see in \Cref{fig:python_multiprocessing}, the energy consumption correlates with the number of processes until the limit of physical cores, when concurrent processes compete for the CPU.

@@

\paragraph*{Synthesis}
This study demonstrates that the optimal way to reduce the energy consumption of Python applications is to use the multiprocessing strategy, which is the most energy-efficient one. However, this strategy should be used cautiously as it can lead to increased execution times.
This study demonstrates that the optimal number of processes is equal to the number of physical cores, which is four in our case.
By using the multithreading strategy, it was also found that sometimes the performance of the application has to be given up in order to save energy.

\clearpage
\input{python_interperters.tex}
% \subsection{threads to validity \note{missing}}
\clearpage
\section{Summary}

Since many software services use Python in public and private cloud infrastructures, making Python-based apps more energy efficient will lower ICT carbon emissions.
This chapter discusses various ways of optimizing the energy consumption of Python-based applications. It first explains why such a choice is relevant. Then it studies the energy behavior of Python within its most commonly employed use cases, which are revealed to be web development and machine learning.

%  TODO : add description about the technology used in the training model 
First, we studied the energy consumption of a machine learning algorithm using the Cifar-10 benchmark to train a model. We discovered that this energy consumption increases exponentially while the model accuracy increases.
As a result, a reduction in accuracy can result in considerable energy savings.
We investigated how data structures, parallelism, and iterative methods affect energy use.

As for web development, we looked at a website built with Django, which is one of the most popular web frameworks in the Python community. We found that fetching the data from the database consumes most of the energy during the request processing phase. Therefore, Django-based web servers can use up to ten times less energy if the developer chooses the write strategy to fetch the data.

After that, we analyzed the impact of several iteration mechanisms in python and their impact on energy usage. We found that the built-in functions and the best practices for writing Python code are the most energy-efficient ones. This energy efficiency is because most of these built-in functions are written in a lower programming language, C.

Finally, we studied the impact of concurrency on energy consumption for Python applications. We found that the multi-processing strategy is the most energy-efficient one. However, this strategy should be used cautiously. This study demonstrates that the optimal number of processes is equal to the number of physical cores, and exceeding this number will cause the application loses its efficiency in both performance and energy consumption. As for the multithreading strategy, We found that the Python interpreter is not thread-safe, which leads to a slower performance than the sequential one. However, this also leads to lower power consumption, which sometimes overcomes the lack of performance to make the application more energy-efficient.

In the second section, we presented a simple technique for optimizing the energy usage of Python-based apps. Without making substantial changes to the code, This is the use of many Python interpreters
Section 2 started by categorizing and filtering these implementations into three significant kinds (compiler, interpreter, and extra libraries).
Then we conducted a series of experiments to compare the energy consumption of these alternatives.
The findings indicate the lack of a general solution and the importance of selecting the appropriate interpreter for the appropriate use case. In certain circumstances, some of these answers may be the best, while others may be the worst.
We found that most interpreters had a similar or worse energy consumption to the official implementation of CPython of their Pythons' version. The reason behind such behavior was that some implementations focused on specific usage cases such as machine learning or security. In contrast, others focused on the compatibility of the python code with their platforms like Jython and IronPython.
Finally, regardless of implementation, we demonstrated that using JIT is the most efficient technique to boost the energy efficiency of Python-based programs.

We believe our contributions will benefit a broad spectrum of legacy systems, reducing ICT's carbon footprint and lowering cloud bills for these services' resources.
